---
title: "Use data that is not IFFI data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Use data that is not IFFI data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ggplot2)
library(purrr)
library(sf)
library(rainfallR)
```


- Many times one wants to extract the rainfall for landslide locations (or non-landslide locations) that might not come from the IFFI database

- The `rainfallR` pacakge was build only having the IFFI data in mind, so there are some quirks that I'll show here

- First, lets create some data



# Make the data

```{r}
# get the shape of ST
st = iffitoR::get_shape_southtyrol() %>% st_transform(crs=32632)

# Get some random points
bb = st_bbox(st)
xs = sample(seq(bb[[1]], bb[[3]], length.out=100), 50)
ys = sample(seq(bb[[2]], bb[[4]], length.out=100), 50)

pts = map2_df(xs, ys, ~data.frame(x = .x, y=.y)) %>% 
  st_as_sf(coords = c("x", "y"), crs=32632)

pts_st = pts[st, ]

# make up some dates
pts_st[["date"]] = sample(seq(as.Date("2000-01-01"), as.Date("2020-01-01"), by="day"), nrow(pts_st))
```


Now that we have some data, lets query the rainfall


```{r, error=TRUE}

r_data = get_rainfall_point_data(
  pts_st,
  days_back = 10,
  daily_thresh = 1.1,
  n_dry = 2,
  ncores = 2,
  save = T,
  base_path = "~/rainfalldata"
)

```


- But uhhh, there is no column called `date_info`. So what does this super non-informative message mean?

- In the iffi-database (actually only in the already derived product) I put a column that informs about the maximum level of date availbale as only slides with day-information should be processed... 

- So one (and the only for now;) solution is to add this column in the following way manually: 

```{r}

# as we have all day information in the synthetic dataset...
pts_st[["date_info"]] = "day"

```


So lets try again to get the rainfall data

```{r, error=T, results=F, cache=T}

r_data = get_rainfall_point_data(
  pts_st,
  days_back = 10,
  daily_thresh = 1.1,
  n_dry = 2,
  ncores = 2,
  save = T,
  base_path = "~/rainfalldata"
)


```

- Ok still does not work....
- The problem is a missing unique identifier.
- Let's add a unique identifier to each slide

```{r, error=T, results=F, cache=T}
pts_st[["id"]] = 1:nrow(pts_st)

r_data = get_rainfall_point_data(
  pts_st,
  days_back = 10,
  daily_thresh = 1.1,
  n_dry = 2,
  ncores = 2,
  save = T,
  base_path = "~/rainfalldata",
  id_landslide = "id"
)

```

- Ok yet another error. Pfisshhh, what the fluck....

- The problem is that the unique identifier should not contain the letters `I`and `D` subsequently in upper or lowercase as this is kind of hardcoded into the `PIFF_ID` variable... 
- So lets give them another unique identifier

```{r, results=F}
pts_st[["lll"]] = 1:nrow(pts_st)

r_data = get_rainfall_point_data(
  pts_st,
  days_back = 10,
  daily_thresh = 1.1,
  n_dry = 2,
  ncores = 2,
  save = T,
  base_path = "~/rainfalldata",
  id_landslide = "lll"
)


```

What, no error...


```{r}
print(head(r_data))
```


So we can plot the precpitation for each of these synthetic slide locations


```{r}
ggplot(r_data) +
  geom_path(
    aes(
      x = days_before_event,
      y = precip,
      group = id
    )
  ) +
  scale_x_reverse() +
  theme_void() +
  labs(
    x = "Days before landslide",
    y = "Daily cumulated precipitation [ mm ]"
  ) +
  theme(
    axis.title = element_text(color="#37323e", size=14, family="Times", face="italic"),
    axis.title.y = element_text(color="#37323e", size=14, family="Times", face="italic", angle=90),
    axis.line.y = element_line(color="#37323e", size=1),
    axis.line.x = element_line(color="#37323e", size=1),
    axis.ticks = element_line(color="black"),
    axis.text = element_text(color="#37323e", family="Times", face="italic", margin = margin(rep(5, 4))),
    plot.margin = unit(rep(2,4), "cm"))
```










