---
title: "Extracting data for bozen"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{example_bozen}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

- read the necessary libraries


```{r setup, echo=T, message=FALSE, warning=FALSE}
library(rainfallR)
library(here)
library(sf)
library(dplyr)
library(tmap)
library(mapview)
library(tidyverse)
library(raster)
library(leaflet)
```

- first we need to set some paths

```{r}
# the path to the directory of the NetCDFs which are stored for each month
path_ncdf = "\\\\projectdata.eurac.edu/projects/Proslide/PREC_GRIDS_updated/"

# then we need a path to a spatial object (shape, geopackage...)
spatial.obj = bozen # should be included in the package
head(spatial.obj)
```

- check to see if we are in the right location

```{r, echo=F}
data(World)
tmap_mode("view")
tm_shape(spatial.obj) +
tm_dots()
```


> set some more options

  + **`dts=c(as.Date("2016-01-12"), as.Date("2016-01-14"))`**: The dates we want the data for
  + **`seqq=TRUE`**: We want all the dates inbetween. So set seqq (sequence) to true
  + **`days_back=4`**: For each date we want to have the rainfall for four days prior to the date we are considering

```{r, echo=T}
dts = c(as.Date("2016-01-12"), as.Date("2016-01-14"))
seqq = T 
days_back = 4
```

> get the rainfall data

- what we will get back is a list of three `sf`-dataframes. One for each day of consideration

```{r getdata, echo=T, warning=F, message=F}
res = get_rainfall(data_path = path_ncdf,
                   spatial.obj = spatial.obj,
                   dts = dts,
                   seqq = seqq, 
                   days_back = days_back)
```


- os one can see, the `days_back`, so the antecedent days are not stored in one column, but each day prior to the day of consideration is an own column. 

- This is not the best way to store the data

- we want to bring the data in "long"-format and additionally calculate the cumulative antecedent rainfall


```{r longformat}
res_long = rainfallR::make_cumulative_rainfall(res)
```


- now we have a list with 3 (one for each date) "long" dataframes and an additional column for the comulative rainfall

- if we would have more than one point, they would have an id an be stored as rows

```{r, message=F}
library(purrr)
# lapply(res_long, head, n=2)
map(res_long, ~head(.x, n=2))
```


## simulate the extration for two points

```{r makepoint}
# create a random point right next to bolzano
bozen2 = bozen 

x = st_coordinates(bozen)[,1] + 0.1
y = st_coordinates(bozen)[,2] + 0.1
#modify the x and y coordinates of the new point
bozen2 = data.frame(x,y) %>% st_as_sf(coords=c(1,2), crs=st_crs(bozen)) 
# bind them together
bozen2 = rbind(bozen, bozen2)
```

- where did the second point land?

```{r}
data(World)
tmap_mode("view")
tm_shape(bozen2) +
tm_dots()
```



```{r res2, warning=F, message=F, warning=F}
res2 = get_rainfall(data_path = path_ncdf,
                   spatial.obj = bozen2,
                   dts = dts,
                   seqq = seqq, 
                   days_back = days_back)
```
- the two points appear as two lines in the output

```{r}
res2
```


- bring it again in long format 
- In order to seperate the points and group by the spatial object, an id-column is now created

```{r id}
res2_long = rainfallR::make_cumulative_rainfall(res2)
```



- we can now plot the data for the one-point and the two(n)-point situation

- **I think it is a rather unusual situation that one wants data for more than one date under consideration. Thus, the creation of the dataframe here below would not be necessary**

```{r gg1, echo=F}

# bind all the dates together and make the name a column to identify

# initalizes an empty df
df_long = res_long[[1]]
df_long$date_consid = names(res_long)[[1]]

for (day in seq_along(res_long)) {
  
  # skip the first one
  if(day!=1){
    df = res_long[[day]]
    df$date_consid = names(res_long)[[day]]
    df_long = rbind(df_long, df)
  }
  
}

ggplot(df_long) +
  geom_col(aes(x=date, y=cumsum)) +
  facet_wrap(~date_consid) +
  theme_light() +
  labs(y="Cumulative Antecedent Rainfall in Bozen")
  
```

- do it for the two points

```{r}
# bind all the dates together and make the name a column to identify

# initalizes an empty df
df_long2 = res2_long[[1]]
df_long2$date_consid = names(res2_long)[[1]]

for (day in seq_along(res2_long)) {
  
  # skip the first one
  if(day!=1){
    df = res2_long[[day]]
    df$date_consid = names(res2_long)[[day]]
    df_long2 = rbind(df_long2, df)
  }
  
}

ggplot(df_long2) +
  geom_col(aes(x=date, y=cumsum, fill=as.factor(id)), position="dodge") +
  facet_wrap(~date_consid) +
  theme_light() +
  labs(y="Cumulative Antecedent Rainfall in Bozen")

```


## get the unterlying rasters to confirm if this makes sense

```{r readraster, warning=F}
# make the path to the raster on the eurac server
path_raster = rainfallR::get_nc_paths(data_path = path_ncdf, day = dts[[1]], days_back = 5)

# select the 12th, 13th and 14th day
days = c(12,13,14)
raster_brick = raster::brick(path_raster)
selected_raster = subset(raster_brick, days)

# transform the stack and the polygon
selected_raster = projectRaster(selected_raster, crs="+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")
bozen2 = st_transform(bozen2, 4326)

# what have we got?
pal = colorNumeric(c("#0C2C84", "#41B6C4", "#FFFFCC"), values(selected_raster[[1]]),
  na.color = "transparent")

bz_sp = as(bozen2, Class="Spatial")

groups_rasters = seq(dts[[1]], dts[[2]], by="day") %>% sapply(., as.character)
p1 = selected_raster[[1]]
p2 = selected_raster[[2]]
p3 = selected_raster[[3]]

leaflet() %>% 
  addTiles() %>% 
  addCircles(data=bozen2, group = "shape") %>% 
  addRasterImage(p1, group=groups_rasters[[1]], colors=pal) %>% 
  addRasterImage(p2, group=groups_rasters[[2]], colors=pal) %>% 
  addRasterImage(p3, group=groups_rasters[[3]], colors=pal) %>% 
  addLayersControl(overlayGroups = c("shape", groups_rasters))


```


