---
title: "Work with NetCDF data in R"
output: rmarkdown::html_vignette
author: Robin
date: " `r Sys.Date()` "
vignette: >
  %\VignetteIndexEntry{extract_sample_raster}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=F, warning=F}
library(rainfallR)
library(iffitoR)
library(tidyverse)
library(scales)
library(stars)
library(ncdf4)
```


# Create the paths to the NetCDF

- First, we set some paths. This assumes that the data is available in exactly the format described on the startpage of the vignette

```{r}
# the path to the netcdf
path_data = "\\\\projectdata.eurac.edu/projects/Proslide/PREC_GRIDS_updated/"

# set the day you want the data
day = as.Date("2016-08-16")

# days back
days_back = 2
```

- now we can generate the path(s) to the NetCDF-files

```{r}
# this will return a list of paths. One for each month of data that we need
paths = rainfallR::get_nc_paths(path_data, day, days_back)
```

# Load the NetCdf and inspect the data

```{r, echo=F}
# open the data as ncdf4 object
ncin = ncdf4::nc_open(paths[[1]])

# read it into a stars object
ncin_stars = read_ncdf(paths[[1]])
```

- now we can inspect the data a bit and see that the data has three dimensions and one attributes

```{r}
ncin_stars
```

- we can also access them programatically

```{r, message=F}
# get the dimensions
dimensions = st_dimensions(ncin_stars)

# get the values of the only attribute
# this is a 3d-array 
values = ncin_stars$precipitation
```

# Get the data for one day

- we need to index the fourth index in order to subset the dates

```{r}
# get the 16th day
single_day = ncin_stars[,,,16]
st_dimensions(single_day)
```

- and then we can plot it

```{r}
# plot(single_day, main="Precipation on the 16th of August 2016")
```

